---
title: "Euromanifesto 2024 Documents"
author: "Giuseppe Carteny, Philipp König, Divya Mani Adhikari"
format:
  pdf:
    pdf-engine: lualatex
    mainfont: "Times New Roman"
---

```{r Admin, echo=FALSE, warning=FALSE, error=FALSE, results='asis',message=FALSE}

want <- c('here', 'magrittr', 'data.table', 'dplyr','tibble', 'tidyr', 
          # 'stringi', 'stringr',
          'openxlsx', 'haven', "countrycode", "glue",
          'ggplot2', "ggthemes", "patchwork", 'ggpubr',
          "kableExtra")
have <- want %in% rownames(installed.packages())
if(any(!have)) {install.packages(want[!have])}
junk <- lapply(want, library, character.only=T)
options(scipen = 99)

rm(list = ls())

summary_spreadsheet <- read.xlsx(here('data', 'em2024_list_20260724.xlsx'))

summary_spreadsheet <-
  summary_spreadsheet |> 
  mutate(
    division_id = ifelse(division_id=='EL', 'GR', division_id),
    countryname = countrycode(division_id, origin = 'iso2c', destination = 'country.name.en') 
  )
  

summary_df <- summary_spreadsheet |> filter(data_for_cdbk==1)

# Need to add countrynames with countrycode package

```


```{r summary_stats, echo=FALSE, results='asis', warning=FALSE, error=FALSE, message=FALSE}

ep_lists <- nrow(summary_df)
doc_n <- sum(summary_df$availability )

doc_perc <- paste0(round((doc_n/ep_lists)*100,0), '%')

man_n <- length(summary_df$doc_type[summary_df$doc_type==1])
other_n <- length(summary_df$doc_type[summary_df$doc_type==2])

man_perc <- paste0(round((man_n/doc_n)*100,0), '%')
other_perc <- paste0(round((other_n/doc_n)*100,0), '%')

summary_votes <- 
  summary_df |> 
  summarise(
    .by = countryname,
    votes_docs = sum(votes_percent[availability==1]),
    votes_man = sum(votes_percent[doc_type==1]),
    votes_noman = sum(votes_percent[doc_type %in% c(2,3)])
  )

missing_countries <- summary_votes |> filter(votes_docs==0) %>% .$countryname
avail_countries_n <- length(unique(summary_votes$countryname)) - length(missing_countries)

summary_votes <- summary_votes |> filter(!countryname %in% missing_countries)

mean_avail_votes <- paste0(round(mean(summary_votes$votes_docs),2),'%')
sd_avail_votes <- paste0(round(sd(summary_votes$votes_docs),2),'%')
max_avail_votes <- paste0(round(max(summary_votes$votes_docs),2),'%')
min_avail_votes <- paste0(round(min(summary_votes$votes_docs),2),'%')
min_avail_votes_country <- summary_votes$countryname[summary_votes$votes_docs==min(summary_votes$votes_docs)]
max_avail_votes_country <- summary_votes$countryname[summary_votes$votes_docs==max(summary_votes$votes_docs)]


top_5_missing <- 
  summary_df |> 
  as_tibble() |> 
  filter(availability==0) |> 
  filter(!countryname %in% missing_countries) |> 
  arrange(-votes_percent) %>% 
  filter(!grepl('Indep', x = party_label)) |> 
  filter(votes_percent>3)

top_5_mis_acronym <- glue::glue_collapse(trimws(unique(top_5_missing$party_acronym)), sep = ", ", last = ", and ")
top_5_mis_country <- glue::glue_collapse(trimws(unique(top_5_missing$countryname)), sep = ", ", last = ", and ")
top_5_mis_vperc <- glue::glue_collapse(paste0(trimws(unique(top_5_missing$votes_percent)),'%'), sep = ", ", last = ", and ")

doc_votes_plot <- 
  summary_votes |> 
  dplyr::select(countryname, votes_noman, votes_man) |> 
  pivot_longer(dplyr::starts_with('votes')) |>
  mutate(name = ifelse(name=='votes_man', 'Manifesto', 'Other')) |> 
  rename(`Document type` = name) |> 
  ggplot(aes(x=reorder(countryname, value), y=value, colour=`Document type`, fill=`Document type`, linetype=`Document type`)) +
  geom_bar(stat='identity') +
  scale_colour_manual(values=c('#0072B2', 'indianred'), aesthetics = c('fill', 'colour')) +
  coord_flip() +
  theme_tufte() +
  ylab('Valid votes (%)') +
  xlab(' ') +
  theme(legend.position = 'top')

doc_avail_plot <- 
  summary_df |> 
  as_tibble() |> 
  mutate(
    .by=division_id, 
    party_number = row_number(),
    total_votes = sum(votes_percent[availability==1])
  ) |> 
  filter(!countryname %in% missing_countries) |> 
  ggplot(aes(x=party_number, y=reorder(countryname, total_votes), colour=as.factor(availability))) +
  geom_point(size=3,shape=15) +
  scale_colour_manual(name='', values=c('#E1BE6A', '#40B0A6'), labels = c('Not available', 'Available')) +
  scale_x_continuous(breaks=seq(1,20,1)) +
  theme_tufte() +
  theme(legend.position = 'top', axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  ylab('') +
  xlab('Party')

availability_comb_plot <- ggarrange(doc_votes_plot,doc_avail_plot)

word_count_mean <- 
  prettyNum(
    round(mean(summary_df$word_count_doc1[summary_df$availability==1]),0),
    big.mark = ","
  )

word_count_sd <- 
  prettyNum(
    round(sd(summary_df$word_count_doc1[summary_df$availability==1]),0),
    big.mark = ","
  )


summary_df <- 
  summary_df |> 
  mutate(
    word_count_mean = round(mean(word_count_doc1[availability==1]),0),
    word_count_sd = round(sd(word_count_doc1[availability==1]),0),
    word_count_mean = prettyNum(word_count_mean, big.mark = ','),
    word_count_sd = prettyNum(word_count_sd, big.mark = ',')
  )

```

# Summary

This document summarises the collection, preprocessing, and systematisation of the party manifestos presented by the electoral lists that participated in the 2024 European Parliament (EP) elections --- hereinafter referred to as *Euromanifestos*. The materials were collected between October 2024 and July 2026 by several research groups, with the process ultimately finalised by the Euromanifesto team of the Societal Observatory Using Novel Data Sources (SOUNDS) project at Saarland University. The collection relied primarily on searches of institutional and party websites and their archived versions, including webpages preserved by the Internet Archive’s Wayback Machine.[^wayback]

# Selection and inclusion criteria

We define electoral programmes or manifestos as official programmatic documents produced by political actors that outline the issues, policy priorities, and proposals promoted in a specific election. Where comparable programmatic documents were not available, we relied on other materials—such as leaflets, press releases, or news articles—published by the relevant electoral lists before the 2024 elections and a sufficiently comprehensive account of their electoral priorities and policy positions. Where available, we relied on official PDF documents produced by the electoral lists that participated in the EP elections. When electoral manifestos were published as webpages, we scraped the relevant content and converted it into PDF documents. In line with the case selection criteria applied by earlier iterations of the Euromanifesto project, we include in our corpus the documents of all the lists that obtained at least one seat in the 2024 EP elections, as recorded in the official results provided by the EP.[^ep_results_database]

# Summary statistics 

We collected documents for `r doc_n` of the `r ep_lists` electoral lists that participated in the 2024 EP elections, and secured at least one seat in the EP, corresponding to approximately `r doc_perc` of the total. Of these `r doc_n`, `r other_n` are not manifestos or electoral programmes. 
We were not able to collect manifestos or other relevant documents for `r length(missing_countries)` `r ifelse(length(missing_countries)>1,"countries", "country")` (`r glue::glue_collapse(unique(missing_countries), sep = ", ", last = ", and ")`). Therefore, our corpus consists of party manifestos from `r avail_countries_n` countries. 
The collected document represent on the average `r mean_avail_votes` of the valid votes (SD: ``r sd_avail_votes`), with a maximum value of `r max_avail_votes` in `r max_avail_votes_country` and a minimum of `r min_avail_votes` in `r min_avail_votes_country`.
The five most relevant missing documents are those for `r top_5_mis_acronym` that obtained `r top_5_mis_vperc` of the valid votes in the respective party systems (`r top_5_mis_country`). 


```{=latex}
\begin{figure}[!ht]
\centering
\caption{Documents coverage by country}
\label{fig:coverage}
```

```{r availability_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=4, dpi=320}
availability_comb_plot
```


```{=latex}
\begin{flushleft}
\footnotesize
\textit{Note}: 
\end{flushleft}
\end{figure}
```

Document length is highly variable. The average lenght of the whole set of documents is equal to `r word_count_mean` words, with a standard deviation equal to `r word_count_sd` words.

[^wayback]: Internet Archive, *Wayback Machine*. <https://web.archive.org/>

[^ep_results_database]: EP elections' electoral data. <https://results.elections.europa.eu/en/tools/download-datasheets/>

