---
title: "Euromanifesto 2024 Documents"
author: "Giuseppe Carteny, Philipp König, Divya Mani Adhikari"
format:
  pdf:
    pdf-engine: lualatex
    mainfont: "Times New Roman"
    
bibliography: cdbk_bib.bib
---

```{r Admin, echo=FALSE, warning=FALSE, error=FALSE, results='asis',message=FALSE}

want <- c('here', 'magrittr', 'data.table', 'dplyr','tibble', 'tidyr', 
          # 'stringi', 'stringr',
          'openxlsx', 'haven', "countrycode", "glue", "kableExtra",
          'ggplot2', "ggthemes", "patchwork", 'ggpubr',
          "kableExtra")
have <- want %in% rownames(installed.packages())
if(any(!have)) {install.packages(want[!have])}
junk <- lapply(want, library, character.only=T)
options(scipen = 99)

rm(list = ls())

summary_spreadsheet <- read.xlsx(here('data', 'em2024_list_v1.xlsx'))

summary_spreadsheet <-
  summary_spreadsheet |> 
  mutate(
    division_id = ifelse(division_id=='EL', 'GR', division_id),
    countryname = countrycode(division_id, origin = 'iso2c', destination = 'country.name.en') 
  )
  

summary_df <- summary_spreadsheet |> filter(data_for_cdbk==1)

# Need to add countrynames with countrycode package

```


```{r summary_stats, echo=FALSE, results='asis', warning=FALSE, error=FALSE, message=FALSE}

ep_lists <- nrow(summary_df)
doc_n <- sum(summary_df$availability )

doc_perc <- paste0(round((doc_n/ep_lists)*100,0), '%')

man_n <- length(summary_df$doc_type[summary_df$doc_type==1])
noman_n <- length(summary_df$doc_type[summary_df$doc_type==2])
other_n <- length(summary_df$doc_type[summary_df$doc_type==3])

man_perc <- paste0(round((man_n/doc_n)*100,0), '%')
noman_perc <- paste0(round((noman_n/doc_n)*100,0), '%')
other_perc <- paste0(round((other_n/doc_n)*100,0), '%')

nocoalitions_n <- length(summary_df$is_coalition[summary_df$availability==1 & summary_df$is_coalition==0])
coalitions_n <- length(summary_df$is_coalition[summary_df$availability==1 & summary_df$is_coalition==1])
coalitions_perc <- paste0(round(coalitions_n/doc_n,2)*100, '%')
nocoalitions_perc <- paste0(100-round(coalitions_n/doc_n,2)*100, '%')



summary_votes <- 
  summary_df |> 
  summarise(
    .by = countryname,
    votes_docs = sum(votes_percentage[availability==1]),
    votes_man = sum(votes_percentage[doc_type==1]),
    votes_noman = sum(votes_percentage[doc_type %in% c(2)]),
    votes_other = sum(votes_percentage[doc_type %in% c(3)])
  )

missing_countries <- summary_votes |> filter(votes_docs==0) %>% .$countryname
avail_countries_n <- length(unique(summary_votes$countryname)) - length(missing_countries)

summary_votes <- summary_votes |> filter(!countryname %in% missing_countries)

mean_avail_votes <- paste0(round(mean(summary_votes$votes_docs),2),'%')
sd_avail_votes <- paste0(round(sd(summary_votes$votes_docs),2),'%')
max_avail_votes <- paste0(round(max(summary_votes$votes_docs),2),'%')
min_avail_votes <- paste0(round(min(summary_votes$votes_docs),2),'%')
min_avail_votes_country <- summary_votes$countryname[summary_votes$votes_docs==min(summary_votes$votes_docs)]
max_avail_votes_country <- summary_votes$countryname[summary_votes$votes_docs==max(summary_votes$votes_docs)]


top_5_missing <- 
  summary_df |> 
  as_tibble() |> 
  filter(availability==0) |> 
  filter(!countryname %in% missing_countries) |> 
  arrange(-votes_percentage) %>% 
  filter(!grepl('Indep', x = party_label)) |> 
  filter(votes_percentage>3)

top_5_mis_acronym <- glue::glue_collapse(trimws(unique(top_5_missing$party_acronym)), sep = ", ", last = ", and ")
top_5_mis_country <- glue::glue_collapse(trimws(unique(top_5_missing$countryname)), sep = ", ", last = ", and ")
top_5_mis_vperc <- glue::glue_collapse(paste0(trimws(unique(top_5_missing$votes_percentage)),'%'), sep = ", ", last = ", and ")


doc_votes_df   <- 
  summary_votes |> 
  dplyr::select(countryname, votes_noman, votes_man, votes_other) |> 
  pivot_longer(dplyr::starts_with('votes')) |>
  mutate(name = case_when(
    name=='votes_man' ~ 'Manifesto',
    name=='votes_noman' ~ 'Pledges or else',
    name=='votes_other' ~ 'Other'
  )) |> 
  rename(`Document type` = name) 


doc_votes_df$`Document type` <- factor(x = doc_votes_df$`Document type`, levels = c('Manifesto', 'Pledges or else', 'Other'))

doc_votes_plot <- 
  doc_votes_df |> 
  ggplot(aes(x=reorder(countryname, value), y=value, colour=`Document type`, fill=`Document type`, linetype=`Document type`)) +
  geom_bar(stat='identity') +
  scale_colour_manual(values=c('#E66100', '#5D3A9B', '#000000'), aesthetics = c('fill', 'colour')) +
  coord_flip() +
  theme_tufte() +
  ylab('Valid votes (%)') +
  xlab(' ') +
  theme(legend.position = 'top')

doc_avail_plot <- 
  summary_df |> 
  as_tibble() |> 
  mutate(
    .by=division_id, 
    party_number = row_number(),
    total_votes = sum(votes_percentage[availability==1])
  ) |> 
  filter(!countryname %in% missing_countries) |> 
  ggplot(aes(x=party_number, y=reorder(countryname, total_votes), colour=as.factor(availability))) +
  geom_point(size=3,shape=15) +
  scale_colour_manual(name='', values=c('#E1BE6A', '#40B0A6'), labels = c('Not available', 'Available')) +
  scale_x_continuous(breaks=seq(1,20,1)) +
  theme_tufte() +
  theme(legend.position = 'top', axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  ylab('') +
  xlab('Party')

availability_comb_plot <- ggarrange(doc_votes_plot,doc_avail_plot)

word_count_mean <- 
  prettyNum(
    round(mean(summary_df$word_count_doc1[summary_df$availability==1]),0),
    big.mark = ","
  )

word_count_sd <- 
  prettyNum(
    round(sd(summary_df$word_count_doc1[summary_df$availability==1]),0),
    big.mark = ","
  )


summary_df <- 
  summary_df |> 
  mutate(
    word_count_mean = round(mean(word_count_doc1[availability==1]),0),
    word_count_sd = round(sd(word_count_doc1[availability==1]),0),
    word_count_mean = prettyNum(word_count_mean, big.mark = ','),
    word_count_sd = prettyNum(word_count_sd, big.mark = ',')
  )

```

# Summary

This document summarises the collection, preprocessing, and systematisation of the party manifestos presented by the electoral lists that participated in the 2024 European Parliament (EP) elections --- hereinafter referred to as *Euromanifestos*. The materials were collected between October 2024 and July 2026 by several research groups, with the process ultimately finalised by the Euromanifesto team of the Societal Observatory Using Novel Data Sources (SOUNDS) project at Saarland University. The collection relied primarily on searches of institutional and party websites and their archived versions, including webpages preserved by the Internet Archive’s Wayback Machine.[^wayback]

# Selection and inclusion criteria

We define electoral programmes or manifestos as official documents produced by political actors that outline the issues, policy priorities and proposals they are promoting in a specific election. Where comparable documents were unavailable, we relied on other materials, such as pledges, leaflets, bullet points and press releases or news articles, which provided sufficient detail on their electoral priorities and policy positions. The `doc_type` variable systematise these differences, and distinguishes programmatic documents by their thematic scope and structure: documents that cover the party's programmatic agenda across policy fields, with thematic chapter structure and programmatic argument, are coded as full manifestos (1); enumerated pledges, leaflets, and bullet-point formats are coded as short programmes (2). Documents that are not programmatic documents at all — press statements and press releases, interviews, rally and press-conference transcripts, candidate pledges, and website captures without programme content — are coded as other material (3).
Where available, we relied on official PDF documents produced by the electoral lists that participated in the EP elections. When electoral manifestos were published as webpages, we scraped the relevant content and converted it into PDF documents. 


We include in our corpus the documents of all the lists that were recorded in the official results provided by the EP for the 2024 elections.[^ep_results_database] We then offer also the vote share (variable ``) and seats obtained by the electoral lists so to provide access to another selection criterion for further analysis.

Party identifiers were assigned based on those provided by the EP (variable `ep_id`) or those of the Euromanifesto project [variables `em_id`, see @carteny_european_2023]. Parties covered in previous waves retained their established identifier if they were matched directly or via legal continuity, or if they were dominant coalition partners. For new electoral lists, we assigned codes that had never been used in previous waves, using the Euromanifesto logic (country code for the first two digits, family code for the third digit, and a new code for the last two digits). 

# Summary statistics 

```{=latex}
\begin{figure}[!ht]
\centering
\caption{Documents coverage by country}
\label{fig:coverage}
```

```{r availability_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=4, dpi=320}
availability_comb_plot
```


```{=latex}
\begin{flushleft}
\footnotesize
\textit{Note}: 
\end{flushleft}
\end{figure}
```

We collected documents for `r doc_n` of the `r ep_lists` electoral lists that participated in the 2024 EP elections and that were recorded in the official EP election results. This corresponds to approximately `r doc_perc` of the total. The vast majority of these documents refer to party lists (`r nocoalitions_n`, `r nocoalitions_perc`) whereas only `r coalitions_n` (`r coalitions_perc`, ) of them are electoral coalition documents. 
Of these `r doc_n`,  `r man_n` (`r man_perc`) are party manifestos or official electoral programmes, `r noman_n` (`r noman_perc`) are non-manifesto electoral campaign materials, and `r other_n` (`r other_perc`) are other kind of relevant documents. 
We were not able to collect manifestos or other relevant documents for `r length(missing_countries)` `r ifelse(length(missing_countries)>1,"countries", "country")` (`r glue::glue_collapse(unique(missing_countries), sep = ", ", last = ", and ")`). Therefore, our corpus consists of party manifestos from `r avail_countries_n` countries. 
The collected document represent on the average `r mean_avail_votes` of the valid votes (SD: `r sd_avail_votes`), with a maximum value of `r max_avail_votes` in `r max_avail_votes_country` and a minimum of `r min_avail_votes` in `r min_avail_votes_country`.
The five most relevant missing documents are those for `r top_5_mis_acronym` that obtained `r top_5_mis_vperc` of the valid votes in the respective party systems (`r top_5_mis_country`).  

Document length is highly variable. The average lenght of the whole set of documents is equal to `r word_count_mean` words, with a standard deviation equal to `r word_count_sd` words. 

[^wayback]: Internet Archive, *Wayback Machine*. <https://web.archive.org/>

[^ep_results_database]: EP elections' electoral data. <https://results.elections.europa.eu/en/tools/download-datasheets/>

```{r}
#| label: tbl-long-table
#| tbl-cap: "Long table"
#| echo: false
#| results: asis

summary_df %>%
  filter(availability == 1) %>%
  dplyr::select(
    countryname,
    ep_id,
    starts_with("em_id"),
    party_acronym,
    party_label
  ) %>%
  rename(
    `Country` = countryname,
    `EP ID` = ep_id,
    `EM ID (1)` = `em_id_1`,
    `EM ID (2)` = `em_id_2`,
    `Acronym` = party_acronym, 
    `Name` = party_label 
  ) |> 
    mutate(
    across(everything(), ~ tidyr::replace_na(as.character(.x), ""))
  ) |> 
  kbl(
    format = "latex",
    booktabs = TRUE,
    longtable = TRUE,
    escape = TRUE,
    align = "l"
  ) %>%
  kable_styling(
    latex_options = "repeat_header",
    font_size = 8
  ) %>%
  column_spec(
    5,
    width = "2.5cm",
    latex_valign = "p"
  ) %>%
  column_spec(
    6,
    width = "6cm",
    latex_valign = "p"
  )
```

# References

