---
title: "Euromanifesto Study 2024 (Documents)"

author:
  - name: Giuseppe Carteny
    affiliations:
      - name: Department of European Social Research, Saarland University
        city: Saarbrücken
        country: Germany

  - name: Philipp König
    affiliations:
      - name: Department of European Social Research, Saarland University
        city: Saarbrücken
        country: Germany

  - name: Divya Mani Adhikari
    affiliations:
      - name: Department of Computer Science, Saarland University
        city: Saarbrücken
        country: Germany

  - name: Ann-Kathrin Reinl
    affiliations:
      - name: GESIS - Leibniz Institute for the Social Sciences
        city: Mannheim
        country: Germany
  
  - name: Rosa Maria Navarrete
    affiliations:
      - name: Department of Political Science and International Relations, Universidad Autonoma de Madrid
        city: Madrid
        country: Spain

  - name: Alexander Hartland
    affiliations:
      - name: Department of European Social Research, Saarland University
        city: Saarbrücken
        country: Germany

  - name: Spencer James Corp
    affiliations:
      - name: Department of Sociology, The University of North Carolina at Chapel Hill
        city: Chapel Hill
        country: United States

  - name: Sayman Johannes Rohleder-Stribl
    affiliations:
      - name: Department of Political Science, The University of North Carolina at Chapel Hill
        city: Chapel Hill
        country: United States

format:
  pdf:
    pdf-engine: lualatex
    mainfont: "Times New Roman"
    template-partials:
      - title.tex
    include-in-header:
      text: |
        \usepackage[noblocks]{authblk}
        \renewcommand\Authfont{\normalsize}
        \renewcommand\Affilfont{\small}
        \renewcommand\Authsep{, }
        \renewcommand\Authands{, }

geometry:
  - left=2.5cm
  - right=2.5cm
  - top=2.5cm
  - bottom=2.5cm

bibliography: cdbk_bib.bib
---

```{r Admin, echo=FALSE, warning=FALSE, error=FALSE, results='asis',message=FALSE}

want <- c('here', 'magrittr', 'data.table', 'dplyr','tibble', 'tidyr', 
          # 'stringi', 'stringr',
          'openxlsx', 'haven', "countrycode", "glue", "kableExtra",
          'ggplot2', "ggthemes", "patchwork", 'ggpubr',
          "kableExtra")
have <- want %in% rownames(installed.packages())
if(any(!have)) {install.packages(want[!have])}
junk <- lapply(want, library, character.only=T)
options(scipen = 99)

rm(list = ls())

summary_spreadsheet <- read.xlsx(here('data', 'em2024_list_v1.xlsx'))

summary_spreadsheet <-
  summary_spreadsheet |> 
  mutate(
    division_id = ifelse(division_id=='EL', 'GR', division_id),
    countryname = countrycode(division_id, origin = 'iso2c', destination = 'country.name.en') 
  )
  

summary_df <- summary_spreadsheet |> filter(data_for_cdbk==1)

# Need to add countrynames with countrycode package

```


```{r summary_stats, echo=FALSE, results='asis', warning=FALSE, error=FALSE, message=FALSE}

ep_lists <- nrow(summary_df)
doc_n <- sum(summary_df$availability )

doc_perc <- paste0(round((doc_n/ep_lists)*100,0), '%')

man_n <- length(summary_df$doc_type[summary_df$doc_type==1])
noman_n <- length(summary_df$doc_type[summary_df$doc_type==2])
other_n <- length(summary_df$doc_type[summary_df$doc_type==3])

man_perc <- paste0(round((man_n/doc_n)*100,0), '%')
noman_perc <- paste0(round((noman_n/doc_n)*100,0), '%')
other_perc <- paste0(round((other_n/doc_n)*100,0), '%')

party_pdf_n <- length(summary_df$source_type[summary_df$availability==1 & summary_df$source_type=='party_pdf'])
scraped_n <- length(summary_df$source_type[summary_df$availability==1 & summary_df$source_type=='scraped_web'])

party_pdf_perc <- paste0((round(party_pdf_n/doc_n,2))*100,'%')
scraped_perc <- paste0((round(scraped_n/doc_n,2))*100,'%')

nocoalitions_n <- length(summary_df$is_coalition[summary_df$availability==1 & summary_df$is_coalition==0])
coalitions_n <- length(summary_df$is_coalition[summary_df$availability==1 & summary_df$is_coalition==1])
coalitions_perc <- paste0(round(coalitions_n/doc_n,2)*100, '%')
nocoalitions_perc <- paste0(100-round(coalitions_n/doc_n,2)*100, '%')



summary_votes <- 
  summary_df |> 
  summarise(
    .by = countryname,
    votes_docs = sum(vs_2024[availability==1]),
    votes_man = sum(vs_2024[doc_type==1]),
    votes_noman = sum(vs_2024[doc_type %in% c(2)]),
    votes_other = sum(vs_2024[doc_type %in% c(3)])
  )

missing_countries <- summary_votes |> filter(votes_docs==0) %>% .$countryname
avail_countries_n <- length(unique(summary_votes$countryname)) - length(missing_countries)

summary_votes <- summary_votes |> filter(!countryname %in% missing_countries)

mean_avail_votes <- paste0(round(mean(summary_votes$votes_docs),2),'%')
sd_avail_votes <- paste0(round(sd(summary_votes$votes_docs),2),'%')
max_avail_votes <- paste0(round(max(summary_votes$votes_docs),2),'%')
min_avail_votes <- paste0(round(min(summary_votes$votes_docs),2),'%')
min_avail_votes_country <- summary_votes$countryname[summary_votes$votes_docs==min(summary_votes$votes_docs)]
max_avail_votes_country <- summary_votes$countryname[summary_votes$votes_docs==max(summary_votes$votes_docs)]


top_5_missing <- 
  summary_df |> 
  as_tibble() |> 
  filter(availability==0) |> 
  filter(!countryname %in% missing_countries) |> 
  arrange(-vs_2024) %>% 
  filter(!grepl('Indep', x = party_label)) |> 
  filter(vs_2024>3)

top_5_mis_acronym <- glue::glue_collapse(trimws(unique(top_5_missing$party_acronym)), sep = ", ", last = ", and ")
top_5_mis_country <- glue::glue_collapse(trimws(unique(top_5_missing$countryname)), sep = ", ", last = ", and ")
top_5_mis_vperc <- glue::glue_collapse(paste0(trimws(unique(top_5_missing$vs_2024)),'%'), sep = ", ", last = ", and ")


doc_votes_df   <- 
  summary_votes |> 
  dplyr::select(countryname, votes_noman, votes_man, votes_other) |> 
  pivot_longer(dplyr::starts_with('votes')) |>
  mutate(name = case_when(
    name=='votes_man' ~ 'Manifesto',
    name=='votes_noman' ~ 'Short programme',
    name=='votes_other' ~ 'Other'
  )) |> 
  rename(`Document type` = name) 


doc_votes_df$`Document type` <- factor(x = doc_votes_df$`Document type`, levels = c('Manifesto', 'Pledges or else', 'Other'))

doc_votes_plot <- 
  doc_votes_df |> 
  ggplot(aes(x=reorder(countryname, value), y=value, colour=`Document type`, fill=`Document type`, linetype=`Document type`)) +
  geom_bar(stat='identity') +
  scale_colour_manual(values=c('#E66100', '#5D3A9B', '#000000'), aesthetics = c('fill', 'colour')) +
  coord_flip() +
  theme_tufte() +
  ylab('Valid votes (%)') +
  xlab(' ') +
  theme(legend.position = 'top')

doc_avail_plot <- 
  summary_df |> 
  as_tibble() |> 
  mutate(
    .by=division_id, 
    party_number = row_number(),
    total_votes = sum(vs_2024[availability==1])
  ) |> 
  filter(!countryname %in% missing_countries) |> 
  ggplot(aes(x=party_number, y=reorder(countryname, total_votes), colour=as.factor(availability))) +
  geom_point(size=3,shape=15) +
  scale_colour_manual(name='', values=c('#E1BE6A', '#40B0A6'), labels = c('Not available', 'Available')) +
  scale_x_continuous(breaks=seq(1,20,1)) +
  theme_tufte() +
  theme(legend.position = 'top', axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  ylab('') +
  xlab('Party')

availability_comb_plot <- ggarrange(doc_votes_plot,doc_avail_plot)

word_count_mean <- 
  prettyNum(
    round(mean(summary_df$word_count_doc1[summary_df$availability==1]),0),
    big.mark = ","
  )

word_count_sd <- 
  prettyNum(
    round(sd(summary_df$word_count_doc1[summary_df$availability==1]),0),
    big.mark = ","
  )

man_word_count_mean <- 
  prettyNum(
    round(mean(summary_df$word_count_doc1[summary_df$availability==1 & summary_df$doc_type==1]),0),
    big.mark = ","
  )

man_word_count_sd <- 
  prettyNum(
    round(sd(summary_df$word_count_doc1[summary_df$availability==1 & summary_df$doc_type==1]),0),
    big.mark = ","
  )


summary_df <- 
  summary_df |> 
  mutate(
    word_count_mean = round(mean(word_count_doc1[availability==1]),0),
    word_count_sd = round(sd(word_count_doc1[availability==1]),0),
    word_count_mean = prettyNum(word_count_mean, big.mark = ','),
    word_count_sd = prettyNum(word_count_sd, big.mark = ',')
  )

```

# Summary

This document summarises the collection, preprocessing, and systematisation of the party manifestos presented by the electoral lists that participated in the 2024 European Parliament (EP) elections --- hereinafter referred to as *Euromanifestos*. The materials were collected between October 2024 and July 2026 by several research groups at Saarland University (Germany), the European University Institute (Italy), and the University of North Carolina at Chapel Hill (United States).[^hiwis] The process was ultimately finalised by the Euromanifesto team of the Societal Observatory Using Novel Data Sources (SOUNDS) project at Saarland University. The collection was based on a systematic review of institutional and party websites and their archived versions, including webpages preserved by the Internet Archive's Wayback Machine.[^wayback] The corpus, its accompanying documentation, and the code used to convert PDF documents and scraped web content into Markdown files are available in a public, version-controlled repository.[^repository]

[^hiwis]: The Saarland team would also like to thank Shahla Jahangiri for her invaluable work and support in the earlier phases of the document collection.
[^repository]: Euromanifesto 2024 repository: <https://github.com/giucarny/em24>.

# Selection and inclusion criteria

Party manifestos are programmatic documents released ahead of an election that put a party's political ideas and goals on record. As authoritative statements of party policy—typically ratified by party conventions—they represent the party as a whole rather than individual members or factions, cover a broad range of themes and policy positions, and thus contain a "set of key central statements" of party preferences [@budge_ideology_1987]. Their pre-electoral and recurrent release renders them uniquely suited to comparative and diachronic analyses of issue emphases and policy positions [@carter_text_2023]. Following the Euromanifesto Project, a Euromanifesto is accordingly understood as a manifesto issued by a political party ahead of an election to the European Parliament.
Where comparable documents were unavailable, we relied on other materials, such as pledges, leaflets, and press releases or news articles, which provided sufficient detail on their electoral priorities and policy positions. The `doc_type` variable systematise these differences, and distinguishes programmatic documents by their thematic scope and structure, differentiating mostly between (1) documents that cover the party's programmatic agenda across policy fields, with thematic chapter structure and programmatic argument (i.e., full manifestos) and (2) enumerated pledges, leaflets, and bullet-point formats, which can be interpreted as short electoral programmes. Documents that are not programmatic documents — press statements and press releases, interviews, rally and press-conference transcripts, single candidate pledges — are coded as other material (3).
Where available, we relied on official PDF documents produced by the electoral lists that participated in the EP elections. When electoral manifestos were published as webpages, we scraped the relevant content and converted it into PDF documents. 


We include in our corpus the documents of all the lists that were recorded in the official results provided by the EP for the 2024 elections.[^ep_results_database] We then offer also the vote share (variable `vs_2024`) and seats obtained by the electoral lists so to provide access to another selection criterion for further analysis.

Party identifiers were assigned based on those provided by the EP (variable `ep_id`) or those of the Euromanifesto project [variable `em_id`, see @carteny_european_2023]. Parties covered in previous waves retained their established identifier if they were matched directly or via legal continuity. For new electoral lists, we assigned codes that had never been used in previous waves, using the Euromanifesto logic (country code for the first two digits, family code for the third digit, and a new code for the last two digits). 
Coalitions retain their historical coalition code where the alliance itself was coded in previous waves; coalitions without a coded predecessor receive a new identifier, regardless of their constituent parties. Whenever it was doubtful that a 2024 coalition is identical to a previously coded one[^portexample] a new identifier was assigned.

# Summary statistics 

```{=latex}
\begin{figure}[!ht]
\centering
\caption{Representativeness of the corpus across countries}
\label{fig:coverage}
```

```{r availability_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=4, dpi=320}
availability_comb_plot
```


```{=latex}
\begin{flushleft}
\footnotesize
\textit{Note}: The left panel reports the cumulative share of valid votes (in percentage points) obtained in the 2024 European Parliament elections by electoral lists available in the corpus, disaggregated by document type. The right panel shows the electoral lists considered in each country (one square per list): teal squares indicate lists for which a document is available, while yellow squares indicate lists for which no document was retrieved. Countries are ordered by the share of valid votes covered by the corpus.
\end{flushleft}
\end{figure}
```

We collected documents for `r doc_n` of the `r ep_lists` electoral lists that participated in the 2024 EP elections and that were recorded in the official EP election results. This corresponds to approximately `r doc_perc` of the total. The vast majority of these documents, `r nocoalitions_n` texts (`r nocoalitions_perc` of the corpus), refer to party lists, whereas only `r coalitions_n` (`r coalitions_perc`) of them are electoral coalition documents. 
Of these `r doc_n`,  `r man_n` (`r man_perc`) are party manifestos or official electoral programmes, `r noman_n` (`r noman_perc`) are non-manifesto electoral campaign materials, and `r other_n` (`r other_perc`) are other kind of relevant documents. Of the whole set of available documents, `r party_pdf_n` (`r party_pdf_perc`) are official documents provided by the electoral lists, while `r scraped_n` (`r scraped_perc`) are materials obtained via web scraping from online sources - tipically, the lists' websites.

We were not able to collect manifestos or other relevant documents for `r length(missing_countries)` `r ifelse(length(missing_countries)>1,"countries", "country")` (`r glue::glue_collapse(unique(missing_countries), sep = " and ")`). Therefore, our corpus consists of party manifestos from `r avail_countries_n` countries. 
The collected documents represent on average `r mean_avail_votes` (SD: `r sd_avail_votes`) of the valid votes obtained by the electoral lists that participated in the EP elections, with a maximum value of `r max_avail_votes` in `r max_avail_votes_country` and a minimum of `r min_avail_votes` in `r min_avail_votes_country` (see Figure \ref{fig:coverage}).
The five most relevant missing documents are those for `r top_5_mis_acronym`, which obtained `r top_5_mis_vperc` of the valid votes in their respective party system (`r top_5_mis_country`).  

Document length is highly variable. The average lenght of the whole set of documents is equal to `r word_count_mean` words, with a standard deviation equal to `r word_count_sd` words. This holds also considering just full manifestos (Mean: `r man_word_count_mean`; SD: `r man_word_count_sd`).



[^portexample]: For instance, AD in Portugal added the PPM to the previously coded PSD–CDS alliance and therefore received a new code rather than the historical one.

[^wayback]: Internet Archive, *Wayback Machine*. <https://web.archive.org/>

[^ep_results_database]: EP elections' electoral data. <https://results.elections.europa.eu/en/tools/download-datasheets/>

# Electoral lists represented in the corpus

```{r}
#| label: tbl-long-table
#| tbl-cap: "List of electoral lists for which a document is available in the corpus"
#| echo: false
#| results: asis

summary_df %>%
  filter(availability == 1) %>%
  dplyr::select(
    countryname,
    ep_id,
    em_id,
    party_acronym,
    party_label
  ) %>%
  rename(
    `Country` = countryname,
    `EP ID` = ep_id,
    `EM ID` = `em_id`,
    `Acronym` = party_acronym, 
    `Name` = party_label 
  ) |> 
    mutate(
    across(everything(), ~ tidyr::replace_na(as.character(.x), ""))
  ) |> 
  kbl(
    format = "latex",
    booktabs = TRUE,
    longtable = TRUE,
    escape = TRUE,
    align = "l",
    linesep = ""
  ) %>%
  kable_styling(
    latex_options = "repeat_header",
    font_size = 8
  ) %>%
  column_spec(
    4,
    width = "2.5cm",
    latex_valign = "p"
  ) %>%
  column_spec(
    5,
    width = "6cm",
    latex_valign = "p"
  )
```

# References

